\documentclass{beamer}

\usepackage[english]{babel}
\usepackage{graphicx,hyperref,ru,url}

\title{Privacy in Cloud Computing}

\author{Aram Verstegen \and Erik Boss}

\institute{Radboud University Nijmegen}

\date{\today}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

% These frametitles are kinda tongue-in-cheek, not sure if I want to keep them -- Erik

\section{Introduction}
% I estimate this section at about 20 minutes (depending on amound questions)

\begin{frame}
    \frametitle{What are we dealing with?}
    /* General definitions, explanations on cloud computing: SaaS/PaaS/IaaS
    (with a picture?), public/private/community/hybrid clouds  */
\end{frame}

\begin{frame}
    \frametitle{Why your manager likes it?}
    /* Why use cloud computing at all; virtualization, scaling, fast
    deployment, pay-for-use, etc  */
\end{frame}

\begin{frame}
    \frametitle{Why Bruce Schneier has his doubts?}
    /* What are the problems; general overview of the security requirements and
    problems (depending on type of cloud); picture(s) about cloud security for
    fun and profit */
\end{frame}

\begin{frame}
    \frametitle{Why your data might not like it?}
    /* Global overview on pecific privacy concerns w.r.t. clouds; example(s)
    of how things can go badly; primarily: why does it matter! */
\end{frame}

\section{Legal aspects}
% Another 20 minutes should be doable

\begin{frame}
    \frametitle{Because a little reminder is always nice}
    /* Quick recap of the most important and relevant legislation, i.e., data
    protection and privacy directives. */
\end{frame}

\begin{frame}
    \frametitle{As always: "it depends"}
    /* Legal analysis differs given the type of cloud computing (so break it
    down); SLAs should be explained here (if not earlier) */
\end{frame}

\begin{frame}
    \frametitle{There is more to law than code}
    /* If we do manage to find relevant case law, put it here */
\end{frame}

\begin{frame}
    \frametitle{Reinventing the wheel}
    /* Some room for discussion: how does our audience feel about the legal
    bits and pieces? */
\end{frame}

\section{Technical aspects}

\begin{frame}
    \frametitle{To trust of not to trust}
    /* In-depth technical privacy in cloud setting; What does privacy mean in
    this context, how can it be achieved (in general): all sorts of things are involved
    (network security, web security, encryption, computing under encryption,
    etc) */
\end{frame}

\begin{frame}
    \frametitle{The best defense is a good\dots}
    /* Attacks, attacks and more attacks; If possible, frighten the audience!
    Malicious images, simple(r) web application vulnerabilities, etc */
\end{frame}

\begin{frame}
    \frametitle{The best offense is a good\dots}
    /* Countermeasures; PETs; best practices; more detailed explanations of
    systems that give privacy guarantees */
\end{frame}

\begin{frame}
    \frametitle{I can already see the ending}
    /* Obligatory 'Questions?' slide, possibly with a funny picture. */
\end{frame}

\end{document}
